Comprehensive study of the global phase diagram of the $J$-$K$-$\Gamma$ model on a triangular lattice

### 2021-01-24 paper/paper.tex
Paper 2021-01-24 14:17

\documentclass[aps,prb,reprint,amsfonts,amsmath,amssymb,showpacs,groupedaddress,superscriptaddress]{revtex4-1}
\usepackage{bm}
\usepackage{color}
\usepackage{graphicx}
\usepackage[colorlinks,urlcolor=blue,linkcolor=blue,anchorcolor=blue,citecolor=blue,bookmarks]{hyperref}
\usepackage{changes}
    \definechangesauthor[name={Shi Wang}, color=red]{swang}
    \setauthormarkup{}

@@

In order to better understand the quantum phases identified by the ED method, we also perform classical Monte Carlo simulations. We begin with paralleling-tempering Monte Carlo~\cite{Hukushima96} on 40 replicas with temperature $T/J$ ranging from $0.001$ to $1.0$. For each replica, we sample it with a combination of heat-bath~\cite{Miyatake84} and over-relaxation method~\cite{Berg} mainly on a $24 \times 24$ triangle lattice with periodic boundary conditions. A whole Monte Carlo step consists of a single heat-bath sweep and subsequent 10 over-relaxation sweeps over the entire lattice. We perform $10^6$ Monte Carlo steps per replica, then, we copy out the spin configuration from the lowest-T replica, and sample it with a combination of zero-temperature heat-bath and over-relaxation method to get the GS. The zero-temperature heat-bath sampling is simply aligning the spins according to their local fields:
\begin{align}
    \vec{S}_{i} = \frac{\vec{h}_{i}^{loc}}{|\vec{h}_{i}^{loc}|} S,
\end{align}
with
\begin{align}
    \vec{h}_{i}^{loc} = \sum_{\langle j \rangle} J \vec{S}_{j} + K {S}^{\gamma}_{j} \hat{e}^{\gamma} + \Gamma (S^{\alpha}_{j} \hat{e}^{\beta} + S^{\beta}_{j} \hat{e}^{\alpha}),
\end{align}
\added[swang]{and the over-relaxation sampling is just mirror reflection of a spin with respect to the local field:
\begin{eqnarray}
    \vec{S}_i^{new} & = &\frac{2 (\vec{S}_i^{old} \cdot \vec{h}_i^{loc}) \vec{h}_i^{loc}}{|\vec{h}_i^{loc}|^2}-\vec{S}_i^{old}.
\end{eqnarray}
}For some competing states, we start from several different initial configurations in order to obtain the correct classical GS.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{QuantumGlobalPhaseDiagram.pdf}
    \caption{\label{fig:QuantumPhaseDiagram}(Color online) \added[swang]{Zero-temperature global} phase diagram of the triangular lattice $J$-$K$-$\Gamma$ model. The angle $\alpha$ and $\beta$ denote the radial and azimuthal angles, respectively. There are in total ten phases including three FM phases denoted as FM-A, FM-B and FM-C, three stripe phases denoted as Stripe-A, Stripe-B and Stripe-C, a modulated stripe phase, a 120$^\circ$ N\'{e}el phase, a dual N\'{e}el phase and a possible QSL phase. Phase boundaries are determined by the singularities of $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ from ED calculations except for that depicted by the dashed line between the modulated stripe and stripe-C phases given by the classical analyses.}
\end{figure*}

The \added[swang]{zero-temperature} quantum phase diagrams obtained by the ED method for $\Gamma \geq 0$ and $\Gamma \leq 0$ on a $4 \times 6$ cluster with the periodic boundary condition are presented in Fig.~\ref{fig:QuantumPhaseDiagram}, where the phase boundaries are determined by the location of singularities in $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ (see Appendix \ref{apx:PhaseBoundary} for details). We find nine magnetically ordered phases and a phase ascribed to be a QSL. The nine ordered magnetic phases consist of three stripe phases (dubbed as stripe-A, strip-B and stripe-C in Fig.~\ref{fig:QuantumPhaseDiagram}), three FM phases (dubbed as FM-A, FM-B and FM-C in Fig.~\ref{fig:QuantumPhaseDiagram}), one 120$^\circ$ N\'{e}el phase, one dual N\'{e}el phase and one modulated stripe phase. In the followings, we will discuss the details on the natures of these phases and the corresponding phase transitions.

@@

\added[swang]{Although the singularities in $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ show that there are phase transitions between stripe-A, stripe-B and stripe-C (see Appendix \ref{apx:PhaseBoundary} for details), the differences between these phases cannot be seen from the SSFs, which show the same characteristic wave vector [Fig.~\ref{fig:StructureFactors}(a), (b), (c) and (d)]. In fact, the most commonly used physcial quantity to distinguish different quantum phases is the order parameter. For the magnetically ordered phases, such as the stripe phases discussed here, the order parameter can be defined as $\bm{m}(\bm{Q})=\frac{1}{N} \sum_{i} \langle \Omega | \bm{S}_{i} | \Omega \rangle e^{i\bm{Q} \cdot \bm{r}_{i}}$, where $|\Omega\rangle$ is the ground state and $\bm{Q}$ is the characteristic wave vector. Thus, both the ordering wave vector and the moment direction are the defining characteristics. The difference between the stripe-A, stripe-B and stripe-C phases lies in that they have different moment directions, which we will discuss in detail later.}

@@

For the other two phases with large areas and mainly located in the $J<0$ region of the phase diagram, we show the SSFs of two representative points with $\alpha=0.3\pi, \beta=1.3\pi$ and $\alpha=0.7\pi, \beta=1.65\pi$ in Fig.~\ref{fig:StructureFactors}(f) and (g), respectively. The SSFs are peaked at $\tilde{\Gamma}$ point of the first BZ, so both phases have FM orders and they are denoted with FM-B and FM-C in Fig.~\ref{fig:QuantumPhaseDiagram}. We note that there are two special points with $\Gamma=\pm1$ ($J=K=0$) in the two FM phases, which are located at the phase boundary in the classical phase diagram~\cite{PhysRevB.92.165108}. Thus, it is worthy to discuss the properties of the pure $\Gamma$ models in details, and we perform an analysis based on the ``basin-hopping" global optimization and classical Monte Carlo methods. For $\Gamma=+1$, apart from the FM GS with the spin orientation lying in the lattice plane, we also found disordered states having the same energy as the FM state (see Appendix~\ref{apx:DegeneratedStates} for details). For $\Gamma=-1$, in addition to the FM order perpendicular to the lattice plane, there are also several noncollinear magnetic orders energetically degenerate with the FM state (see Appendix~\ref{apx:DegeneratedStates} for details). However, according to our ED calculations, for the quantum model, quantum fluctuations select the FM ordered states as the GS out of the degenerate manifolds of classical states. This selection of states among the degenerate classical GSs is the so called order-by-disorder mechanism, which has previously been applied in a number of insulating magnets, especially in cases where frustration leads to a degenerate manifold of classical GS configurations that is broken by quantum fluctuations~\cite{JPSJ.54.4494,PhysRevLett.62.2056,PhysRevLett.88.067203,PhysRevB.81.214419,PhysRevLett.109.077204}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{StructureFactors.pdf}
    \caption{\label{fig:StructureFactors}(Color online) SSFs from the ED calculations for representative interaction parameters in different phases: (a) stripe-B, (b) and (c) stripe-A, (d) stripe-C, (e) modulated stripe, (f) FM-B, (g) FM-C, (h) QSL. The green dashed lines mark the first BZ of the triangular lattice. (i) Characteristic wave vectors for the \added[swang]{FM (blue circle), stripe (orange square), 120$^\circ$ N\'{e}el (red up-triangle) and dual N\'{e}el (purple down-triangle)} phases. See Appendix~\ref{apx:SSF} for the SSFs of the 120$^\circ$ N\'{e}el phase, dual N\'{e}el phase and pure $\Gamma$ models.}
\end{figure}

@@

\begin{acknowledgments}
This work was supported by the National Natural Science Foundation of China (Grants No. 11674158, No. 11774152, No. 11774300, \added[swang]{No. 12004191 and No. 12074175}, National Key Projects for Research and Development of China (Grant No. 2016YFA0300401) \added[swang]{and Natural Science Foundation of Jiangsu Province, China (Grant No. BK20200738)}.
\end{acknowledgments}

@@

\section{\label{apx:PhaseBoundary}Phase boundaries in the \added[swang]{zero-temperature} quantum phase diagram}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{SecondDerivatives46.pdf}
    \caption{\label{fig:SecondDerivatives46} (Color online) GS energies and their second derivatives versus $\alpha$ or $\beta$ for six representative paths in the phase diagram. \added[swang]{The ED calculations are performed on a $4 \times 6$ cluster with periodic boundary condition.} The blue lines are the GS energies and the red lines are the second derivatives. In (a)-(c) $\alpha$ is fixed to $0.05\pi$, $0.3\pi$ and $0.75\pi$, respectively. In (d)-(f), $\beta$ is fixed to $0$, $0.5\pi$ and $1.5\pi$, respectively.}
\end{figure}

\added[swang]{
\section{\label{apx:SizeEffect}Effects of system size on the phase boundaries}
To study the effects of system size on the zero-temperature quantum phase diagram, we also carry out ED calculations on a $2 \times 6$ cluster with periodic boundary condition, see Fig.~\ref{fig:SecondDerivatives26}(g) for the shape of the cluster as well as the form of the boundary condition. This cluster and the boundary condition are compatible with most of the magnetic orders, such as FM, stripe, 120$^\circ$ N\'{e}el, dual N\'{e}el and six-chain modulated stripe shown in Fig.~\ref{fig:ModulatedStripe}(a).

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{SecondDerivatives26.pdf}
  \includegraphics[width=\columnwidth]{Cluster26.pdf}
  \added[swang]{
  \caption{\label{fig:SecondDerivatives26} (Color online) GS energies and their second derivatives versus $\alpha$ or $\beta$ along the same paths shown in Fig.~\ref{fig:SecondDerivatives46}. The blue lines are the GS energies and the red lines are the second derivatives. In (a)-(c) $\alpha$ is fixed to $0.05\pi$, $0.2\pi$ and $0.75\pi$, respectively. In (d)-(f), $\beta$ is fixed to $0$, $0.5\pi$ and $1.5\pi$, respectively. (g) The $2 \times 6$ cluster as well as the periodic boundary condition that is used to perform the ED calculations.}
  }
\end{figure}
}

>>>>

\documentclass[aps,prb,reprint,amsfonts,amsmath,amssymb,showpacs,groupedaddress,superscriptaddress]{revtex4-1}
\usepackage{bm}
\usepackage{color}
\usepackage{graphicx}
\usepackage[colorlinks,urlcolor=blue,linkcolor=blue,anchorcolor=blue,citecolor=blue,bookmarks]{hyperref}

@@

In order to better understand the quantum phases identified by the ED method, we also perform classical Monte Carlo simulations. We begin with paralleling-tempering Monte Carlo~\cite{Hukushima96} on 40 replicas with temperature $T/J$ ranging from $0.001$ to $1.0$. For each replica, we sample it with a combination of heat-bath~\cite{Miyatake84} and over-relaxation method~\cite{Berg} mainly on a $24 \times 24$ triangle lattice with periodic boundary conditions. A whole Monte Carlo step consists of a single heat-bath sweep and subsequent 10 over-relaxation sweeps over the entire lattice. We perform $10^6$ Monte Carlo steps per replica, then, we copy out the spin configuration from the lowest-T replica, and sample it with a combination of zero-temperature heat-bath and over-relaxation method to get the GS. The zero-temperature heat-bath sampling is simply aligning the spins according to their local fields:
\begin{align}
    \vec{S}_{i} = \frac{\vec{h}_{i}^{loc}}{|\vec{h}_{i}^{loc}|} S,
\end{align}
with
\begin{align}
    \vec{h}_{i}^{loc} = \sum_{\langle j \rangle} J \vec{S}_{j} + K {S}^{\gamma}_{j} \hat{e}^{\gamma} + \Gamma (S^{\alpha}_{j} \hat{e}^{\beta} + S^{\beta}_{j} \hat{e}^{\alpha}),
\end{align}
{\color{red}and the over-relaxation sampling is just mirror reflection of a spin with respect to the local field:
\begin{eqnarray}
    \vec{S}_i^{new} & = &\frac{2 (\vec{S}_i^{old} \cdot \vec{h}_i^{loc}) \vec{h}_i^{loc}}{|\vec{h}_i^{loc}|^2}-\vec{S}_i^{old}.
\end{eqnarray}
}For some competing states, we start from several different initial configurations in order to obtain the correct classical GS.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.95\textwidth]{QuantumGlobalPhaseDiagram.pdf}
    \caption{\label{fig:QuantumPhaseDiagram}(Color online){ \color{red}Zero-temperature global} phase diagram of the triangular lattice $J$-$K$-$\Gamma$ model. The angle $\alpha$ and $\beta$ denote the radial and azimuthal angles, respectively. There are in total ten phases including three FM phases denoted as FM-A, FM-B and FM-C, three stripe phases denoted as Stripe-A, Stripe-B and Stripe-C, a modulated stripe phase, a 120$^\circ$ N\'{e}el phase, a dual N\'{e}el phase and a possible QSL phase. Phase boundaries are determined by the singularities of $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ from ED calculations except for that depicted by the dashed line between the modulated stripe and stripe-C phases given by the classical analyses.}
\end{figure*}

The {\color{red}zero-temperature} quantum phase diagrams obtained by the ED method for $\Gamma \geq 0$ and $\Gamma \leq 0$ on a $4 \times 6$ cluster with the periodic boundary condition are presented in Fig.~\ref{fig:QuantumPhaseDiagram}, where the phase boundaries are determined by the location of singularities in $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ (see Appendix \ref{apx:PhaseBoundary} for details). We find nine magnetically ordered phases and a phase ascribed to be a QSL. The nine ordered magnetic phases consist of three stripe phases (dubbed as stripe-A, strip-B and stripe-C in Fig.~\ref{fig:QuantumPhaseDiagram}), three FM phases (dubbed as FM-A, FM-B and FM-C in Fig.~\ref{fig:QuantumPhaseDiagram}), one 120$^\circ$ N\'{e}el phase, one dual N\'{e}el phase and one modulated stripe phase. In the followings, we will discuss the details on the natures of these phases and the corresponding phase transitions.

@@

{\color{red}Although the singularities in $-\partial^2E_0/\partial\alpha^2$ and $-\partial^2E_0/\partial\beta^2$ show that there are phase transitions between stripe-A, stripe-B and stripe-C (see Appendix \ref{apx:PhaseBoundary} for details), the differences between these phases cannot be seen from the SSFs, which show the same characteristic wave vector [Fig.~\ref{fig:StructureFactors}(a), (b), (c) and (d)]. In fact, the most commonly used physcial quantity to distinguish different quantum phases is the order parameter. For the magnetically ordered phases, such as the stripe phases discussed here, the order parameter can be defined as $\bm{m}(\bm{Q})=\frac{1}{N} \sum_{i} \langle \Omega | \bm{S}_{i} | \Omega \rangle e^{i\bm{Q} \cdot \bm{r}_{i}}$, where $|\Omega\rangle$ is the ground state and $\bm{Q}$ is the characteristic wave vector. Thus, both the ordering wave vector and the moment direction are the defining characteristics. The difference between the stripe-A, stripe-B and stripe-C phases lies in that they have different moment directions, which we will discuss in detail later.}

@@

For the other two phases with large areas and mainly located in the $J<0$ region of the phase diagram, we show the SSFs of two representative points with $\alpha=0.3\pi, \beta=1.3\pi$ and $\alpha=0.7\pi, \beta=1.65\pi$ in Fig.~\ref{fig:StructureFactors}(f) and (g), respectively. The SSFs are peaked at $\tilde{\Gamma}$ point of the first BZ, so both phases have FM orders and they are denoted with FM-B and FM-C in Fig.~\ref{fig:QuantumPhaseDiagram}. We note that there are two special points with $\Gamma=\pm1$ ($J=K=0$) in the two FM phases, which are located at the phase boundary in the classical phase diagram~\cite{PhysRevB.92.165108}. Thus, it is worthy to discuss the properties of the pure $\Gamma$ models in details, and we perform an analysis based on the ``basin-hopping" global optimization and classical Monte Carlo methods. For $\Gamma=+1$, apart from the FM GS with the spin orientation lying in the lattice plane, we also found disordered states having the same energy as the FM state (see Appendix~\ref{apx:DegeneratedStates} for details). For $\Gamma=-1$, in addition to the FM order perpendicular to the lattice plane, there are also several noncollinear magnetic orders energetically degenerate with the FM state (see Appendix~\ref{apx:DegeneratedStates} for details). However, according to our ED calculations, for the quantum model, quantum fluctuations select the FM ordered states as the GS out of the degenerate manifolds of classical states. This selection of states among the degenerate classical GSs is the so called order-by-disorder mechanism, which has previously been applied in a number of insulating magnets, especially in cases where frustration leads to a degenerate manifold of classical GS configurations that is broken by quantum fluctuations~\cite{JPSJ.54.4494,PhysRevLett.62.2056,PhysRevLett.88.067203,PhysRevB.81.214419,PhysRevLett.109.077204}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{StructureFactors.pdf}
    \caption{\label{fig:StructureFactors}(Color online) SSFs from the ED calculations for representative interaction parameters in different phases: (a) stripe-B, (b) and (c) stripe-A, (d) stripe-C, (e) modulated stripe, (f) FM-B, (g) FM-C, (h) QSL. The green dashed lines mark the first BZ of the triangular lattice. (i) Characteristic wave vectors for the {\color{red}FM (blue circle), stripe (orange square), 120$^\circ$ N\'{e}el (red up-triangle) and dual N\'{e}el (purple down-triangle)} phases. See Appendix~\ref{apx:SSF} for the SSFs of the 120$^\circ$ N\'{e}el phase, dual N\'{e}el phase and pure $\Gamma$ models.}
\end{figure}

@@

\begin{acknowledgments}
This work was supported by the National Natural Science Foundation of China (Grants No. 11674158, No. 11774152, No. 11774300, {\color{red}No. 12004191 and No. 12074175}, National Key Projects for Research and Development of China (Grant No. 2016YFA0300401) {\color{red}and Natural Science Foundation of Jiangsu Province, China (Grant No. BK20200738)}.
\end{acknowledgments}

@@

\section{\label{apx:PhaseBoundary}Phase boundaries in the {\color{red}zero-temperature} quantum phase diagram}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{SecondDerivatives46.pdf}
    \caption{\label{fig:SecondDerivatives46} (Color online) GS energies and their second derivatives versus $\alpha$ or $\beta$ for six representative paths in the phase diagram. {\color{red}The ED calculations are performed on a $4 \times 6$ cluster with periodic boundary condition.} The blue lines are the GS energies and the red lines are the second derivatives. In (a)-(c) $\alpha$ is fixed to $0.05\pi$, $0.3\pi$ and $0.75\pi$, respectively. In (d)-(f), $\beta$ is fixed to $0$, $0.5\pi$ and $1.5\pi$, respectively.}
\end{figure}

{\color{red}
\section{\label{apx:SizeEffect}Effects of system size on the phase boundaries}
To study the effects of system size on the zero-temperature quantum phase diagram, we also carry out ED calculations on a $2 \times 6$ cluster with periodic boundary condition, see Fig.~\ref{fig:SecondDerivatives26}(g) for the shape of the cluster as well as the form of the boundary condition. This cluster and the boundary condition are compatible with most of the magnetic orders, such as FM, stripe, 120$^\circ$ N\'{e}el, dual N\'{e}el and six-chain modulated stripe shown in Fig.~\ref{fig:ModulatedStripe}(a).

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{SecondDerivatives26.pdf}
  {\color{red}
  \caption{\label{fig:SecondDerivatives26} (Color online) GS energies and their second derivatives versus $\alpha$ or $\beta$ along the same paths shown in Fig.~\ref{fig:SecondDerivatives46}. The blue lines are the GS energies and the red lines are the second derivatives. In (a)-(c) $\alpha$ is fixed to $0.05\pi$, $0.2\pi$ and $0.75\pi$, respectively. In (d)-(f), $\beta$ is fixed to $0$, $0.5\pi$ and $1.5\pi$, respectively. (g) The $2 \times 6$ cluster as well as the periodic boundary condition that is used to perform the ED calculations.}
  }
\end{figure}
}
